---
title: "Analysis_model_building"
author: "Wendel Raymond"
date: "October 1, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Model building
Using the data prepared from the data preparation script we can now beging model building. Before we can run a PERMANOVA we will need to prep the response and explanatory data. After that we will generate a global model and eventually best model, testing the effect of habitat on fish community composition.

```{r data}
dat <- read.csv("Data/SEAK_EventID_mass.csv", header = TRUE, stringsAsFactors = TRUE)
```

```{r libraries}
library(dplyr)
library(tidyr)
library(vegan)
```

## Prepping response and explanatory data
For ease of processing later we will separate the response data (species and mass) from the explanatory data (EventID, date, month, temp, salinity, etc.)

```{r separating}
names(dat)
site <- unique(dat[, 2:17])
duplicated(site$EventID)

fish <- dat[, c(1, 2, 18)]
duplicated(fish)
fish <- spread(fish, SpCode, mass_g)
```

