---
title: "Analysis_data_preparation"
author: "Wendel Raymond"
date: "September 25, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data preparation
Before model building the NOAA fish atlas data will need to be prepared. This includes the following steps.

1. Subsetting master data for southeast Alaska
    + Rename habitat types so they are consistent
2. Remove invertebrates
3. Assign lengths to unmeasured fish
4. Convert lengths to mass

## Data import
Import master data. Data were exported from a query of the master database from NOAA. 

```{r data}
dat <- read.csv("Data/NOAA_All_Data_24092018.csv", header = TRUE, stringsAsFactors = FALSE)
lw <- read.csv("Data/SEAK_fish_tax_LW.csv", header = TRUE, stringsAsFactors = FALSE)
```

```{r libraries}
library(dplyr)
library(data.table)
```

## Subsetting and cleanin master data
Southeast Alaska only, habitat name changes for accuracy.

```{r subsetting and cleaning}
## Rename ##
dat$Habitat <- as.factor(dat$Habitat)
levels(dat$Habitat)[levels(dat$Habitat)=="Sand-gravel"] <- "Sand-Gravel"
levels(dat$Habitat)[levels(dat$Habitat)=="Eelgrass"] <- "Seagrass"

## Subset ##
dat.se <- dat %>% 
  filter(Region == "southeastern Alaska") %>% # SE Alaska only
  filter(Habitat != "Surfgrass") %>% # remove stange surfgrass site
  filter(Locale != "Yakutat Bay" & Locale != "Situk Estuary")
```

## Defining invertebrates
Lots of species that were recorded are not fish, and we are not interested in them for this analysis.

```{r define inverts}
## Species names ##
sort(unique(dat.se$SpCode))

## New column that assigns Taxa ##
dat.se$Taxon <- ifelse(dat.se$SpCode == "AMPHIPOD" | dat.se$SpCode == "BSNAILSA"| dat.se$SpCode == "BSNAILW"| dat.se$SpCode == "CRABRR"| dat.se$SpCode == "CRABWIDE"| dat.se$SpCode =="CRABCKE"| dat.se$SpCode == "CRABDEC"| dat.se$SpCode == "CRABDUN"| dat.se$SpCode == "CRABGD"| dat.se$SpCode == "CRABGH"| dat.se$SpCode == "CRABGKE"| dat.se$SpCode == "CRABGRA"| dat.se$SpCode == "CRABHEL" | dat.se$SpCode == "CRABHER"| dat.se$SpCode ==  "CRABLYR"| dat.se$SpCode == "CRABNKE"| dat.se$SpCode == "CRABPRH"| dat.se$SpCode == "DORIDCL"| dat.se$SpCode == "ISOPODC"| dat.se$SpCode == "ISOPODU"| dat.se$SpCode == "JLYCLNG"| dat.se$SpCode == "JLYCOMB"| dat.se$SpCode == "JLYICE"| dat.se$SpCode == "JLYLION"| dat.se$SpCode == "JLYMOON"| dat.se$SpCode == "MUSSELB"| dat.se$SpCode == "MYSID"| dat.se$SpCode == "MYSIDOP"| dat.se$SpCode == "none"| dat.se$SpCode == "NUDHOOD"| dat.se$SpCode == "NUDOPAL"| dat.se$SpCode ==  "NUDRM"| dat.se$SpCode == "OCTOPUS"| dat.se$SpCode == "POLYCHT"| dat.se$SpCode == "SHRMBLD"| dat.se$SpCode ==  "SHRMCOO"| dat.se$SpCode ==  "SHRMCR"| dat.se$SpCode ==   "SHRMDB"| dat.se$SpCode ==   "SHRMDOC"| dat.se$SpCode ==  "SHRMGRA"| dat.se$SpCode ==  "SHRMHEP"| dat.se$SpCode ==  "SHRMHIP"| dat.se$SpCode ==  "SHRMNC"| dat.se$SpCode == "SHRMROU"| dat.se$SpCode ==  "SHRMSID"| dat.se$SpCode ==  "SHRMSK"| dat.se$SpCode ==  "SHRMSPO"| dat.se$SpCode ==  "SHRMSTC"| dat.se$SpCode ==  "SHRMSTI"| dat.se$SpCode ==  "SHRMSTP"| dat.se$SpCode ==  "SHRMTAN"| dat.se$SpCode == "SNAILBB"| dat.se$SpCode == "SNAILBT"| dat.se$SpCode == "SNAILFO"| dat.se$SpCode == "SNAILLH"| dat.se$SpCode ==  "SNAILLV"| dat.se$SpCode ==  "SNAILMM"| dat.se$SpCode ==  "SNAILMP"| dat.se$SpCode ==  "SNAILRT"| dat.se$SpCode == "STAR6ARM"| dat.se$SpCode == "STARLEA"| dat.se$SpCode == "STARMOT"| dat.se$SpCode == "STARPYCN" | dat.se$SpCode == "TUNICAT"| dat.se$SpCode == "UNCLAM"| dat.se$SpCode == "UNISOPOD"| dat.se$SpCode == "UNJLY"| dat.se$SpCode == "UNKRILL"| dat.se$SpCode == "UNMYSID"| dat.se$SpCode == "UNNUD"| dat.se$SpCode == "UNSHRM"| dat.se$SpCode == "UNSNAIL"| dat.se$SpCode == "UNWRM" | dat.se$SpCode == "UNCHGRN"| dat.se$SpCode == "URCHIN"| dat.se$SpCode == "WRMSCA", "INVERT", "VERT")
```

Now that inters are defined they can be filtered out of the data
```{r remove inverts}
dat.se <- dat.se %>% 
  filter(Taxon == "VERT")
```

## Select gear type
A few different seine types were used in the project. We only want data from the 37m variable mesh beach seine. The is designated at BSEINE in the Gear column
```{r gear type}
dat.se <- dat.se %>%
  filter(Gear == "BSEINE")
```

## Assigning lengths to unmeasured fish
Since only the first 50 fish of a given species in a given set were measured and the rest counted, the counted fish need to be assigned lengths based off the lenght-frequency of the measured fish. Note that the below code was adapted from code developed for the APECS beach seine data.

### Separate measured and unmeasured fish
The first step is to separate the measured fish from the unmeasured fish. Note that this removes a TUBEPOA that had neigher a ForLength or an Unmeasured count.
```{r measured - unmeasured split}
## Measured ##
dat.se.m <- dat.se %>%
  filter(ForkLength > 0)

## Unmeasured ##
dat.se.um <- dat.se %>%
  filter(Unmeasured > 0)
  
dat.se.um$Unmeasured <- as.numeric(dat.se.um$Unmeasured)
```

## Assign lengths to unmeasured fish
When beach seining we only measured the first 50 individuals of a species, and counted the rest. We can use the measured fishes to create a distribution from which we can assign lengths to the unmeasured fishes.

Assign lengths to unmeasured fish based on sampled distribution. This assignment should happen at the EventID level. i.e. use the distribution of fishes at a EventID to assign unmeasured fishes at that EventID. 

First we need to filter out data were a fish was counted but there is no corresponding measurment
```{r remove no measures}
d <- data.frame() # empty dataframe to fill with for loop

for(e in unique(dat.se.um$EventID)){ # cycle through uniqie EventIDs
  dat.m <- dat.se.m %>% # subset measured data by interation EventID
    filter(EventID == e)
  dat.um <- dat.se.um %>% # subset unmeasured data by iteration EventID
      filter(EventID == e)
  dat.um.new <- dat.um[dat.um$SpCode %in% dat.m$SpCode, ]
  d <- rbind(d, dat.um.new)
}
```

We will assume the least and just use the sampled proportions to assign lenghts to unmeasured fish. Here, instead of using the raw uneasured data frame I will use the subset data frame made from the loop above that removed instances where an unmeasured fish did not have a measured counterparts.  
```{r assignment direct}
q <- data.frame() # empty dataframe to fill with for loop


for(e in unique(dat.se.um$EventID)){ # cycle through uniqie EventIDs
  dat.m <- dat.se.m %>% # subset measured data by interation EventID
    filter(EventID == e)
  dat.um <- d %>% # subset unmeasured data by iteration EventID
      filter(EventID == e)
  for(i in unique(dat.um$SpCode)){ # cycle through species that are in UNMEASURED data
    samp <- dat.m %>% # create sample from which to make distrubution
      filter(SpCode == i)
    unmeas <- dat.um %>% # isolate unmeasured fish
      filter(SpCode == i)
    unmeas <- as.numeric(unmeas$Unmeasured) # save unmeasured value
    dat.temp1 <- data.frame(size = as.character(samp$ForkLength))
    dat.temp2 <- dat.temp1 %>% 
      group_by(size) %>% 
      summarise(count = n())
    dat.temp2$porb <- (dat.temp2$count/sum(dat.temp2$count))
    dat.temp2$x <- as.numeric(paste(dat.temp2$size))
    fx <- function(n){ # function derived from limits and probabilities of above
    sample(x = min(dat.temp2$x):max(dat.temp2$x), n, replace = TRUE, prob = dat.temp2$prob)
    }
    dat.temp3 <- data.frame(EventID = e, SpCode = i, ForkLength = fx(unmeas))
    q <- rbind(q, dat.temp3) # append iteration to full data
  }
} # this is returning an error in regards to prob, but i am not sure what its talking about
```

Append assigned lengths to master data with all site level data
```{r append assigned lengths}
## Merge with original measured fishes ##
## Extract site and sp data ##
fish.site <- unique(dat.se[,1:31])

## Merge with loop output, the lengths of unmeasured fishes ##
d.info <- merge(q, fish.site, by = "EventID")

## Merge with original measured fishes ##
fish.all <- bind_rows(dat.se.m, d.info)
```

## Append Taxonomy
To help with later data analysis we will append family and order information so grouping by higher taxonomic values can be done.

```{r tax}
## Append Taxonomy ##
fish.all <- merge(fish.all, lw[, 1:5], by = "SpCode")
```


## Calculate Biomass
Using the length-weight conversion values individual lengths will be converted to biomass. Coefficients are in cm*g so fork lengths will need to be converted to cm from mm. Also 

```{r calc mass}
## Unique species in master data ##
sp <- data.frame(SpCode = unique(fish.all$SpCode))

## Species that need stand ins ##
nomatch <- anti_join(sp, lw, by = "SpCode")

## Species that don't have a or b balues ##
nodat <- subset(lw, is.na(lw$a)) 
nodat$SpCommon

# Calculate mean a and b values by Family #
fams <- lw %>% 
  group_by(Family) %>% 
  summarise(n = n(),
            mean_a = mean(a, na.rm = TRUE),
            mean_b = mean(b, na.rm = TRUE))

## Calculate mean a and b values by Species with appended Family ##
sps <- lw %>% 
 group_by(SpCode) %>%
 summarise(Family = unique(Family),
           n = n(),
           mean_a = mean(a, na.rm = TRUE),
           mean_b = mean(b, na.rm = TRUE)) 
```


