---
title: "Data_exploration"
author: "Wendel Raymond"
date: "September 20, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Data Exploration
Examining the NOAA fish atlas data.

```{r libraries}
library(dplyr)
library(DT)
library(ggplot2)
library(leaflet)
library(mapview)
library(sp)
```


```{r data}
noaa <- read.csv("Data/NOAA_All_Data_24092018.csv", header = TRUE, stringsAsFactors = FALSE)
```

## Data structure
```{r structure}
str(noaa)
unique(noaa$EventID)
unique(noaa$Region)
unique(noaa$Exposure)
```

## Data coverage
Since I am interested in habitat effects on fish community composition it is important to understand how many sites are of each hanitat in differnent years and in different locations. But first I need to so some quick data clean up. To rename habitat types and exclude the one surfgrass site.

```{r data clean}
# Rename Sand-gravel to Sand-Gravel
noaa$Habitat <- as.factor(noaa$Habitat)
levels(noaa$Habitat)[levels(noaa$Habitat)=="Sand-gravel"] <- "Sand-Gravel"
```


```{r coverage}
se.hab <- noaa %>% 
  filter(Region == "southeastern Alaska") %>% 
  filter(Habitat != "Surfgrass") %>% 
  group_by(Habitat, Mon) %>% 
  summarise(count = n_distinct(EventID))

se.loc <- noaa %>% 
  filter(Region == "southeastern Alaska") %>% 
  filter(Habitat != "Surfgrass") %>% 
  group_by(Habitat, Location) %>% 
  summarise(count = n_distinct(EventID))
```

So it appears that there is reasonable habitat replication across months, but not some much acorss locations. Is may be necessary to groups spatially at a higher level. For example, how its broken up in the actual atlas migh be best; north inside/outside and south inside/outside. I may also be worth considering adding an exposure factor too. This may help explain things. However, exposure is going to be confounded with habitat type. 

```{r map}
pal <- colorFactor(c("black", "green", "chocolate4", "orange", "blue"), domain = c("Bedrock", "Eelgrass", "Kelp", "Sand-Gravel", "Surfgrass"))

leaflet() %>% 
  addProviderTiles("OpenStreetMap") %>% 
  addMouseCoordinates() %>% 
  addCircleMarkers(data = noaa[noaa$Region == "southeastern Alaska",], ~Long1, ~Lat1, color = ~Location, fillOpacity = 1, stroke = FALSE, radius = 5, label = ~Location) 
```

It appears tht yes, the data should be spatially grouped. Proposed groups are how its broken up in the actual atlas migh be best; north inside/outside and south inside/outside. Yes there are seines in Yakutat but there is not replication of all habitat types.

```{r new region define}
## Make Polygons ##
# North - Outside #
NO.x <- c(-137.09838, -136.06567, -135.96680, -135.69214, -135.03296, -135.04395, -135.76904, -137.30713)
NO.y <- c(58.41898, 58.39020, 57.82136, 57.43312, 57.14816, 56.87900, 56.63811, 58.25173)

NO <- cbind(NO.x, NO.y)
NO.s = Polygon(NO)
NO.sp = Polygons(list(NO.s),1)
NO.sp = SpatialPolygons(list(NO.sp))
```

