---
title: Project overview - Understanding fish communities across nearshore habitats
  in southeast Alaska
author: "Wendel Raymond"
date: "September 7, 2018"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r libraries, warning = FALSE, message = FALSE}
library(dplyr)
library(DT)
library(ggplot2)
```


## Goals and Objectives
- Goal
    + Understand fish community composition across nearshore habitats in southeast Alaska

- Objectives
  1. Build a statistical model to relate community structure (multivariate) to habitat characteristics
  2. Build a statistical model to relate juvenile abundance of harvested species/groups (e.g. Salmonidae, Gadidae) to habitat characteristics

## Background - Data
- Atlas of Nearshore Fishes of Alaska
    + 1998 - 2011
    + 555 unique sites
    + Walleye pollock, Pacific herring, pink salmon, and chum salmon accounted for 55% of total catch

- Habitats
    + Bedrock
    + Eelgrass
    + Understory kelp
    + sand / gravel

## Background - Analyses
- Previous Analyses
    + NOAA technical memoranda and a few peer-reviewed papers 
    + Summary statistics, small geographic scale
    + Fish community appears to differ across habitats

- Data analysis in community ecology class project
    + Began to tackle this question of community composition as a function of habitat
    + Results later

## Data
```{r data}
noaa <- read.csv("Data/NOAA_All_Data.csv", header = TRUE, stringsAsFactors = FALSE)
```

```{r}
names(noaa)
```

