---
title: Project overview - Understanding fish communities across nearshore habitats
  in southeast Alaska
author: "Wendel Raymond"
date: "September 7, 2018"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r libraries, warning = FALSE, message = FALSE}
library(dplyr)
library(DT)
library(ggplot2)
library(leaflet)
```


## Goals and Objectives
- Goal
    + Understand fish community composition across nearshore habitats in southeast Alaska

- Objectives
  1. Build a statistical model to relate community structure (multivariate) to habitat characteristics
  2. Build a statistical model to relate juvenile abundance of harvested species/groups (e.g. Salmonidae, Gadidae) to habitat characteristics

## Background - Data
- Atlas of Nearshore Fishes of Alaska
    + 1998 - 2011
    + 555 unique sites
    + Walleye pollock, Pacific herring, pink salmon, and chum salmon accounted for 55% of total catch

- Habitats
    + Bedrock
    + Eelgrass (Seagrass)
    + Understory kelp
    + Sand / Gravel

## Background - Analyses
- Previous Analyses
    + NOAA technical memoranda and a few peer-reviewed papers 
    + Summary statistics, small geographic scale
    + Fish community appears to differ across habitats

- Data analysis in community ecology class project
    + Began to tackle this question of community composition as a function of habitat
    + Results later

## Data
- Explanatory
    + Site information
    + Lat/Long
    + Temp/Salinity
    + Date
    + Habitat Type

## Data
- Response 
    + Species
    + Life Stage
    + Fork Length
    + Count of unmeasured fish

```{r load data}
noaa <- read.csv("Data/NOAA_All_Data.csv", header = TRUE, stringsAsFactors = FALSE)

## Re order so explanatory variables are on the left and response variables are on the right
noaa <- cbind.data.frame(noaa[,1:10], noaa[15:26], noaa[,11:14])

#####----- Habitat Clean up -----#####
# Rename Sand-gravel to Sand-Gravel
noaa$Habitat <- as.factor(noaa$Habitat)
levels(noaa$Habitat)[levels(noaa$Habitat)=="Sand-gravel"] <- "Sand-Gravel"
levels(noaa$Habitat)[levels(noaa$Habitat)=="Surfgrass"] <- "Eelgrass"
```

## Lets take a look
```{r}
datatable(head(noaa))
```

## Where is the data?
```{r map}
pal <- colorFactor(c("black", "green", "chocolate4", "orange", "blue"), domain = c("Bedrock", "Eelgrass", "Kelp", "Sand-Gravel", "Surfgrass"))
leaflet() %>% 
  addTiles() %>% 
  addCircleMarkers(data = noaa[noaa$Region == "southeastern Alaska",], ~Long1, ~Lat1, color = ~pal(Habitat), fillOpacity = 1, stroke = FALSE, radius = 5, label = ~Habitat) %>% 
  addLegend(pal = pal, values = c("Kelp", "Eelgrass", "Surfgrass", "Bedrock", "Sand-Gravel"), opacity = 1)
```

## Proposed data processing
The data *may* need some work before its ready for full analysis.

1. Assigning unmeasured fish FL based off the population sample
2. Converting FL to mass
3. Append/have full taxonomic information ready for each species

## Previous modeling approaches
Tested for the effect of Habitat on fish community composition 
- Habitat only
- Counts

PERMANOVA results indicated that fish community composition was significantly different among habitats (psudo-F = 17.04, p-value < 0.001). 

## Modeling approach
1. Build a statistical model to relate community structure (multivariate) to habitat characteristics

$ fish comm = Habitat $